### 2019-04-16 report/final-report.tex

The average number of examples to
infer the correct lens does not tell the whole story.  The system will
stop as soon as it finds a well typed lens that satisfies the supplied examples.
This inferred lens may or may not 
correctly handle unseen examples that correspond to
unexercised portions of the source and target regular expressions.
Figure~\ref{fig:exs-reqd} lists
the number of examples that are required to determinize the generation of
permutations in \RigidSynth{}.
Intuitively, this number represents the maximum number of
examples that a user must supply to guide the synthesis engine if it
always guesses the wrong permutation when multiple permutations can be used to
satisfy the specification.

The average number of examples is so much lower than the maximum
number of required examples because of correspondences in how we wrote
the regular expressions for the source and target data formats. 
Specifically, when we had corresponding disjunctions in both the
source and the target, we ordered them the same way.  The algorithm
uses the supplied ordering to guide its search, and so the system
requires fewer examples.   We did not write the examples in this style
to facilitate synthesis, but rather because maintaining similar
subparts in similar orderings makes the types much easier to 
read. We expect that most users would do the same.

@@

For each problem in the benchmark suite, we calculate the following measures
as proxies for the level of programmer effort when using each the three
approaches: \bcp{This seems very detailed, and it includes some technical
  terms (like ``canonize(q)'') that readers of this summary may not
  know.  But maybe it's OK?}

%
\begin{itemize}
  \item[\QRESize{}:] 
  The number of AST nodes in the QRE specifications for the source and
  target formats, including examples. 
  \item[\canonizeAndSpecSize{}:] 
  The sum of (1) the number of AST nodes in $W(q)$ for each QRE $q$ in the source and target
  formats, (2) the number of AST nodes in $\canonize(q)$ for each QRE $q$ with a
  non-trivial canonizer, and (3) the number of AST nodes in the
  examples.  We use (1) to estimate the burden of describing
  the external source and target formats and (2) to estimate the
  burden of writing the requisite canonizers
  by hand.  We count the nodes in the examples because they would be
  fed to the bijective synthesizer.  
  These counts are an approximation, as both $W(q)$ and $\canonize(q)$ are
  automatically generated from the corresponding QRE $q$, and it is
  possible that a human-written version might be smaller.
  \item[\LensAndSpecSize{}:] The sum of (1) the number of AST nodes in
  $W(q)$ for each QRE $q$ in the source and target formats and (2) the
  number of AST nodes in the synthesized QRE lens.  We use (1) to
  estimate the burdern of describing the source and target formats
  and (2) to estimate the burdern of writing the appropriate lens by
  hand. These counts are also approximations, as
  $W(q)$ and the synthesized lens may be larger than one written by hand.
\end{itemize}

>>>>

The average number of examples to
infer the correct lens does not tell the whole story.  The system will
stop as soon as it finds a well typed lens that satisfies the supplied examples.
This inferred lens may or may not 
correctly handle unseen examples that correspond to
unexercised portions of the source and target regular expressions.
We also calculated
the number of examples that are required to determinize the generation of
permutations in \RigidSynth{}.
Intuitively, this number represents the maximum number of
examples that a user must supply to guide the synthesis engine if it
always guesses the wrong permutation when multiple permutations can be used to
satisfy the specification.

The average number of examples needed is much lower than the maximum number of
required examples because of correspondences in how we wrote the regular
expressions for the source and target data formats.  Specifically, when we had
corresponding disjunctions in both the source and the target, we ordered them
the same way.  The algorithm uses the supplied ordering to guide its search, and
so the system requires fewer examples.  We did not write the examples in this
style to facilitate synthesis, but rather because maintaining similar subparts
in similar orderings makes the types much easier to read. We expect that most
users would do the same.

@@

For each problem in the benchmark suite, we calculate the following measures
as proxies for the level of programmer effort when using each the three
approaches:

%
\begin{itemize}
  \item[\QRESize{}:] 
  The number of AST nodes in the QRE specifications for the source and
  target formats, including examples. 
  \item[\canonizeAndSpecSize{}:] 
  The sum of (1) the number of AST nodes in $W(q)$ for each QRE $q$ in the source and target
  formats, (2) the number of AST nodes in $\canonize(q)$ for each QRE $q$
  requiring a  non-trivial canonizer (``canonizers'' as transformations that
  project a string to its equivalence class representative; they are written as
  programs in Boomerang), and (3) the number of AST nodes in the
  examples.  We use (1) to estimate the burden of describing
  the external source and target formats and (2) to estimate the
  burden of writing the requisite canonizers
  by hand.  We count the nodes in the examples because they would be
  fed to the bijective synthesizer.  
  These counts are an approximation, as both $W(q)$ and $\canonize(q)$ are
  automatically generated from the corresponding QRE $q$, and it is
  possible that a human-written version might be smaller.
  \item[\LensAndSpecSize{}:] The sum of (1) the number of AST nodes in
  $W(q)$ for each QRE $q$ in the source and target formats and (2) the
  number of AST nodes in the synthesized QRE lens.  We use (1) to
  estimate the burdern of describing the source and target formats
  and (2) to estimate the burdern of writing the appropriate lens by
  hand. These counts are also approximations, as
  $W(q)$ and the synthesized lens may be larger than one written by hand.
\end{itemize}
